The Supervisionary proof-checking kernel

### 2021-10-25 paper/prisc-2022.tex
prisc-2022: minor grammatical changes

Though ostensibly very different, squinting, we see the two kinds of kernel are tasked with solving the same task: enforcing system invariants in the face of interaction with untrusted code.
Yet, the two solutions to solving this problem, employed by the respective kinds of kernel, are very different.
In this abstract, we explore designing proof-checking kernels as \emph{supervisory software}, where separation between kernel and untrusted code is enforced by \emph{privilege}.

@@

\emph{Supervisionary} is a proof-checking system for Gordon's HOL, structured as supervisory software, and provides a \emph{system interface} to untrusted code similar to an operating system's system call interface.
Supervisionary is implemented as a WebAssembly (Wasm henceforth) host, allowing us to rapidly prototype.

@@

Existing tools in the HOL familiy could be refactored around Supervisionary by changing their kernels to act as a shim between untrusted automation and the Supervisionary kernel.
Arguably, this increases the robustness of existing tools, enforcing separation by isolation, rather than module boundaries.
More interestingly, this also means that Supervisionary acts as a mechanism for ``transporting'' definitions and theorems between systems within the wider HOL family: systems are caapble of referring to, and manipulating, kernel objects produced by other systems through Supervisionary kernel handles.
With this, we can also bootstrap a full theorem-proving environment, with associated libraries of content, on top of Supervisionary without writing it from scratch.

@@

For example, the operational semantics and instruction decoding functionality of Wasm is clearly embeddable in HOL.
Using this, properties of a program $P$ to be executed under Supervisionary could be established statically, and registered with the kernel, perhaps interactively by a user or by another program executing before $P$ executes.
This theorem can then be used by $P$ in closing challenges from Supervisionary, issued at runtime, tieing the correctness of $P$ to its execution.

>>>>

Though ostensibly very different, squinting, we see that the two kinds of kernel are tasked with solving the same task: enforcing system invariants in the face of interaction with untrusted code.
Yet, the two solutions to solving this problem, employed by the respective kinds of kernel, are very different.
In this abstract, we explore designing proof-checking kernels as \emph{supervisory software}, where separation between kernel and untrusted code is enforced by \emph{privilege}.

@@

\emph{Supervisionary} is a proof-checking system for Gordon's HOL, structured as supervisory software, and provides a \emph{system interface} to untrusted code similar to an operating system's system call interface.
Supervisionary is implemented as a WebAssembly (Wasm henceforth) host, allowing us to prototype rapidly.

@@

Existing tools in the HOL familiy could be refactored around Supervisionary by changing their kernels to act as a shim between untrusted automation and the Supervisionary kernel.
Arguably, this increases the robustness of existing tools, enforcing separation by isolation, rather than module boundaries.
More interestingly, this also means that Supervisionary acts as a mechanism for ``transporting'' definitions and theorems between systems within the wider HOL family: systems are capable of referring to, and manipulating, kernel objects produced by other systems through Supervisionary kernel handles.
With this, we can also bootstrap a full theorem-proving environment, with associated libraries of content, on top of Supervisionary without writing it from scratch.

@@

For example, the operational semantics and instruction decoding functionality of Wasm is clearly embeddable in HOL.
Using this, properties of a program $P$ to be executed under Supervisionary could be established statically, and registered with the kernel, perhaps interactively by a user or by another program executing before $P$ executes.
This theorem can then be used by $P$ in closing challenges from Supervisionary, issued at runtime, tying the correctness of $P$ to its execution.
